New lower bounds for Schur and weak Schur numbers

### 2021-05-07 article/weakschur.tex
relecture (pas encore terminée partie weak schur)

First, put the integers of \([\![1, ap + q]\!]\) in the following table (with \(a\) columns and \(p + 1\) lines) and divide it into 3 blocks (the columns are numbered from \(-l\) to \(+q\)):

@@

\begin{definition}
Let \( (a,n,b) \in (\mathbb{N}^*)^3\), Let \((A_1,...,A_n)\) a partition of  \([\![1, a + b]\!]\).
This partition is said to be a \(b\)-weakly-sum-free template (\(b\)-WSF-template) with width \(a\) and \(n\) colors if:

\begin{itemize}
\item \(\forall i \in [\![1, n]\!], \quad A_i\) is weakly-sum-free
\item \(\forall i \in [\![1, n]\!], \quad A_i\backslash [\![1, b]\!]\) is sum-free
\item For \(A_n\) (the special subset): \[\forall (x,y) \in A_n^2, \,
x+y>b+2a \implies x+y-2a\notin A_n\]
\item For the others subsets: \\ \[\forall i \in [\![0,n-1]\!], \, \forall(x,y) \in A_i^2, \,
x+y>a+b \implies \pi(x+y) \notin A_i
\]
\end{itemize}
\end{definition}

@@

\begin{definition}
Let \( (n,b) \in (\mathbb{N}^*)^2\). If there exists \(a\) such that there exists a \(b\)-WSF-template with width \(a\) 
and \(n\) colors, define:
\[
\WS_b^+(n)=-b+\max \{a \in \mathbb{N}^*/ \text{there exists a } b \text{-WSF-template with width } a \text{ and } n \text{ colors} \}
\]
If no such \(a\) exists, set \(\WS_b^+(n) = 0\).
\end{definition}

@@

\begin{theorem}
Let \((a,n,b) \in (\mathbb{N}^*)^3\), let \( (p,k) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,p]\!]\) 
into \(k\) sum-free subsets and a \(b\)-WSF-template \((A_1,...,A_n)\) with width \(a\) and \(n\) colors, 
then there exists a partition of \([\![1, b+p \times a]\!]\) into \((k+n-1)\) weakly sum-free subsets.
\end{theorem}

@@

Define \( f : [\![1,b+pa]\!] \longrightarrow [\![1,n]\!] \) as follows:

@@

We have to verify that f induces weakly-sum-free templates:\\
\\
\underline{if \((x,y) \in \mathcal{T}^2\) such that \(f(x)=f(y)\), \(x \neq y\):}\\
\\\(x+y<a+b\) since \(b<a\) and \(g(x)=f(x)=f(y)=g(y)\).
\\Hence \(f(x+y)=g(x+y)\neq g(x)=f(x)\)
\\\\
\underline{if \((x,y) \in \mathcal{T} \times \mathcal{C}\) such that \(f(x)=f(y)\), \(x \neq y\):}\\
We distinguish two cases:

@@

\underline{if \((x,y) \in \mathcal{T} \times \mathcal{R}\) such that \(f(x)=f(y)\), \(x \neq y\):}\\
\\Then, \(f(x)=f(y)=n\). We distinguish two cases:

@@

\underline{if \((x,y) \in \mathcal{C}^2\) such that \(f(x)=f(y)\), \(x \neq y\):}\\
\\Then  \(g(\pi(x))=f(x)=f(y)=g(\pi(y))\). We distinguish two cases:

@@

\underline{if \((x,y) \in (\mathcal{R})^2\) such that \(f(x)=f(y)\), \(x \neq y\):}\\
\\Let \(r(x)=\pi(x)-a\) and \(r(y)=\pi(y)-a\),
\\We proved that \(x=a\lambda(x)+\pi(x)-a\), therefore \(x=a\lambda(x)+\pi(x)\)
\\\(x+y=a(\lambda(x)+\lambda(y))+r(x)+r(y)\). We distinguish three cases:

@@

\begin{proposition}
Let \((b, k, a) \in \mathbb{N}^*)^3\) and let \(f\) be a coloring associated to a \(b\)-WSF-template with width \(p\) and \(k\) colors. Let
\(c \in \mathbb{N}\) (\(c = \min (A_{k+1} \backslash [\![1, b ]\!]) - 1\) works) and assume there there exists a coloring \(g\) of
\([\![b + 1, b + c]\!]\) with \(k\) colors such that:

@@

And the associated inequality :

@@

\qquad Having found suitable templates, which can be found in the appendix, with a SAT solver, we claim that for all \(n \in
\mathbb{N}^*\):

>>>>

First, put the integers of \([\![1, ap + q]\!]\) in the following table (with \(a\) columns and \(p + 1\) lines) 
and divide it into 3 blocks (the columns are numbered from \(-l\) to \(+q\)).

@@

\begin{definition}
Let \( (a,n,b) \in (\mathbb{N}^*)^3\), Let \((A_1,...,A_n)\) a partition of  \([\![1, a + b]\!]\).
This partition is said to be a \(b\)-weakly-sum-free template (\(b\)-WSF-template) with width \(a\) and \(n\) colors if

\begin{itemize}
\item \(\forall i \in [\![1, n]\!], \quad A_i\) is weakly-sum-free
\item \(\forall i \in [\![1, n]\!], \quad A_i\backslash [\![1, b]\!]\) is sum-free
\item For \(A_n\) (the special subset)
	\[
	\forall (x,y) \in A_n^2, \,x+y>b+2a \implies x+y-2a\notin A_n
	\]
\item For the others subsets
	\[
	\forall i \in [\![0,n-1]\!], \, \forall(x,y) \in A_i^2, \, x+y>a+b \implies \pi(x+y) \notin A_i
	\]
\end{itemize}
\end{definition}

@@

\begin{definition}
Let \( (n,b) \in (\mathbb{N}^*)^2\). If there exists \(a\) such that there exists a \(b\)-WSF-template with width \(a\) 
and \(n\) colors, define
\[
\WS_b^+(n)=-b+\max \{a \in \mathbb{N}^*/ \text{there exists a } b \text{-WSF-template with width } a \text{ and } n \text{ colors} \}
\]
If no such \(a\) exists, set \(\WS_b^+(n) = 0\).
\end{definition}

@@

\begin{theorem}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) and \( (p,k) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,p]\!]\) 
into \(k\) sum-free subsets and a \(b\)-WSF-template \((A_1,...,A_n)\) with width \(a\) and \(n\) colors, 
then there exists a partition of \([\![1, b+p \times a]\!]\) into \((k+n-1)\) weakly sum-free subsets.
\end{theorem}

@@

Define \( f : [\![1,b+pa]\!] \longrightarrow [\![1,n]\!] \) as follows

@@

We have to verify that f induces weakly-sum-free templates\\
\\
\underline{if \((x,y) \in \mathcal{T}^2\) such that \(f(x)=f(y)\), \(x \neq y\):}\\
\\\(x+y<a+b\) since \(b<a\) and \(g(x)=f(x)=f(y)=g(y)\).
\\Hence \(f(x+y)=g(x+y)\neq g(x)=f(x)\)
\\\\
\underline{if \((x,y) \in \mathcal{T} \times \mathcal{C}\) such that \(f(x)=f(y)\), \(x \neq y\):}\\
We distinguish two cases

@@

\underline{if \((x,y) \in \mathcal{T} \times \mathcal{R}\) such that \(f(x)=f(y)\), \(x \neq y\):}\\
\\Then, \(f(x)=f(y)=n\). We distinguish two cases

@@

\underline{if \((x,y) \in \mathcal{C}^2\) such that \(f(x)=f(y)\), \(x \neq y\):}\\
\\Then  \(g(\pi(x))=f(x)=f(y)=g(\pi(y))\). We distinguish two cases

@@

\underline{if \((x,y) \in (\mathcal{R})^2\) such that \(f(x)=f(y)\), \(x \neq y\):}\\
\\Let \(r(x)=\pi(x)-a\) and \(r(y)=\pi(y)-a\),
\\We proved that \(x=a\lambda(x)+\pi(x)-a\), therefore \(x=a\lambda(x)+\pi(x)\)
\\\(x+y=a(\lambda(x)+\lambda(y))+r(x)+r(y)\). We distinguish three cases

@@

\begin{proposition}
Let \((b, k, a) \in (\mathbb{N}^*)^3\) and let \(f\) be a coloring associated to a \(b\)-WSF-template with width \(p\) and \(k\) colors. Let
\(c \in \mathbb{N}\) (\(c = \min (A_{k+1} \backslash [\![1, b ]\!]) - 1\) works) and assume there there exists a coloring \(g\) of
\([\![b + 1, b + c]\!]\) with \(k\) colors such that

@@

And the associated inequality

@@

\qquad Having found suitable templates, which can be found in the appendix, with a SAT solver, we claim that for all \(n \in
\mathbb{N}^*\)
